% ─── CHAPTER 1 ────────────────────────────────────────────────────────────────
\chapter{Introduction}\label{INTRO}

\section{Motivation and Background}\label{INTRO_MOT}

Climate change is one of the defining challenges of the twenty-first
century~\cite{unesco_sdgs}. Understanding historical climate patterns ---
temperature trends, precipitation cycles, and the differences between regions
--- is an essential first step for researchers, educators, policymakers, and
the general public alike. Yet access to high-quality climate data remains, in
practice, restricted to those with specialist GIS or programming skills.
WorldClim~\cite{fick2017worldclim} is a globally comprehensive open dataset
that provides monthly climate and meteorological rasters spanning the period
1951--2024, covering variables such as temperature, precipitation, solar
radiation, and wind speed at resolutions down to 30 arc-seconds
($\approx$900\,m). Despite this richness, querying WorldClim traditionally
requires desktop GIS software, custom scripts, or academic subscriptions ---
placing it out of reach for the majority of potential users.

At the same time, the web platform has matured to the point where rich,
interactive data exploration can be delivered entirely in the browser, with no
installation and no registration. Modern JavaScript frameworks, vector-tile
mapping libraries, and reactive data-fetching primitives make it possible to
build tools that would previously have required a dedicated server-side
pipeline. The GSIC research group at the Universidad de Valladolid has developed a
Virtual Semantic API~\cite{scrapi} that exposes its raster archive as a
SPARQL-compatible REST service, opening the door to exactly this kind of
lightweight, client-only integration.

Despite these advances, existing web tools that offer climate data fall short
in several important ways. Consider Climate-Charts.net~\cite{climatecharts} as
a representative example: it covers only locations with nearby meteorological
stations, meaning large parts of Africa, Central Asia, and the Amazon basin are
poorly represented; it exposes raw station observations rather than
spatially interpolated raster surfaces, so values between stations are
unreliable; its temporal scope is limited to a single historical baseline; and
it provides no mechanism for regional spatial analysis. Similar limitations
affect the other tools surveyed in Chapter~\ref{REL}: narrow regional coverage,
coarse or point-based resolution, restricted time periods, and a static
single-location view with no comparative or temporal exploration.

WorldClim avoids all of these problems --- it provides globally consistent,
interpolated raster data at kilometre-scale resolution across seven decades ---
but exposing such a complex, high-dimensional dataset through an accessible
web interface is itself a significant challenge. No general-purpose,
browser-based application currently allows a non-specialist user to look up a
city by name, inspect its monthly climate profile, compare it to another city,
explore how its climate has shifted across historical periods, and visualise the
spatial distribution of a climate variable across an entire region --- all
within a single, cohesive interface. In this Bachelor Thesis, I propose a web 
application, named \textbf{Climatica}, to fill that gap.

\section{Objectives}\label{INTRO_OBJ}

The primary objective of this Bachelor Thesis is to design, implement, and
evaluate a web application that makes the historical climatic and meteorological
dataset of WorldClim accessible to non-specialist users, allowing them to
freely choose which climate or meteorological raster to explore --- without
requiring registration, local software installation, or GIS expertise.

The specific objectives are as follows:

\begin{enumerate}
  \item \textbf{City climate profile.} Enable a user to search for any city in
    the world by name and view a complete monthly climate profile, including
    minimum temperature, maximum temperature, average temperature, and
    precipitation, together with a Walter-Lieth climatogram~\cite{walter1960} 
    for rapid climate classification.

  \item \textbf{City comparison.} Allow side-by-side comparison of the climate
    profiles of two cities, with overlaid charts that make seasonal differences
    immediately visible.

  \item \textbf{Period comparison.} Allow the user to compare the climate of a
    single city across time, using two distinct modes. In \emph{climate mode},
    the user selects any two of the five available 30-year climate normals
    (1951--1980, 1961--1990, 1970--2000, 1981--2010, and 1991--2020) and views
    their monthly profiles overlaid on shared axes with trend indicator cards.
    In \emph{weather mode}, the user selects up to five individual years from
    the CRU-TS-4.09 weather archive (1951--2024) and compares them
    simultaneously on a single chart.

  \item \textbf{Web architecture.} Implement the application as a client-side
    single-page application~\cite{react} that communicates directly with the
    WorldClim Virtual Semantic API~\cite{scrapi} and the Wikidata~\cite{wikidata}
    knowledge base for city search and geocoding, without developing a custom
    back-end.

  \item \textbf{Internationalisation.} Deliver the interface in at least two
    languages, with city names displayed in the active language.
\end{enumerate}

\section{Methodology}\label{INTRO_MET}

The project followed an iterative, prototype-driven development process
organised into four broad phases, consistent with established software
engineering research methodology~\cite{adrion1993}.

\textbf{Phase 1 --- Research and scoping.} The first two weeks were devoted to
understanding the WorldClim Virtual Semantic API and the Wikidata knowledge
graph, surveying existing climate visualisation tools (see
Chapter~\ref{REL}), and agreeing the functional scope with the supervisors.
User stories were collected and translated into a set of functional and
non-functional requirements (Section~\ref{APP_REQ}).

\textbf{Phase 2 --- Architecture and technology selection.} Based on the
requirements, the technology stack was chosen (Section~\ref{APP_TECH}) and
the application architecture was defined (Section~\ref{APP_ARCH}). The key
architectural decision --- to build a front-end-only application that calls
the WorldClim and Wikidata APIs directly --- was validated against the
WorldClim API documentation and confirmed with the supervisors.

\textbf{Phase 3 --- Iterative implementation.} The four application pages were
implemented one at a time, starting with the City Climate page as it exercised
the core API integration and chart rendering. Each page was demonstrated to the
supervisors and other researchers in the group during our meetings, feedbacks 
was collected, and the next iteration began. This phase constituted the majority 
of the 400 working hours.

\textbf{Phase 4 --- Evaluation and documentation.} A usability evaluation was
conducted with representative users (Section~\ref{EVAL}), the results were
analysed, and this report was written.

Throughout all phases the codebase was managed with Git and hosted on GitHub,
with a linear commit history that maps directly to the phases above.

\section{Structure of the Document}\label{INTRO_EST}

The remainder of this report is organised as follows.

Chapter~\ref{REL} reviews related work, covering existing climate data portals
and visualisation tools as well as the two external services on which Climatica
depends most heavily: the WorldClim dataset and its Virtual Semantic API, and
the Wikidata knowledge graph.

Chapter~\ref{APP} describes the application itself. It begins with the
functional and non-functional requirements, continues with the data model and
architectural decisions, details the technology choices and their rationale, and
concludes with a walkthrough of the four implemented pages.

Chapter~\ref{EVAL} presents the evaluation. It covers the demo scenarios used
to exercise the application, a usability study conducted with target users, and
a discussion of limitations and potential improvements.

Chapter~\ref{CONC} summarises the conclusions, reflects on whether the project
objectives were met, and outlines directions for future work.