% * ─── CHAPTER 5 ────────────────────────────────────────────────────────────────
\chapter{Conclusions}\label{CONC}

This chapter summarises the work carried out during this Bachelor's thesis +
internship project, reflects on whether the original objectives were met, and
identifies directions for future development.

% * ─── 5.1 Summary of Contributions ────────────────────────────────────────────
\section{Summary of Contributions}\label{CONC_SUMMARY}

This Bachelor's thesis describes the design, implementation, and evaluation of
\textbf{Climatica}, a browser-based web application that makes the WorldClim
historical climate and meteorological dataset accessible to non-specialist users
without requiring GIS software, programming skills, or registration. The
application is publicly available at \url{https://climatica.gsic.uva.es/}.

The project addressed a clear gap in the landscape of existing climate data
tools. As the survey in Chapter~\ref{REL} demonstrated, no publicly available
tool combines ease of use, open kilometre-scale raster data, interactive
multi-view exploration, and support for temporal trend analysis in a single
registration-free browser interface. Climatica fills this gap by exposing the
full WorldClim archive --- covering the period 1951--2024 across five 30-year
climate normals and individual weather years --- through four complementary
pages: a city climate profile with a Walter-Lieth climatogram, a side-by-side
city comparison, a multi-period climate change explorer, and a regional heatmap
renderer.

The key technical contributions of the project are as follows:

\begin{itemize}
  \item \textbf{Full-stack Next.js architecture.} The original React SPA was
    redesigned around Next.js~16 with server-side route handlers, moving all
    SCRAPI authentication credentials off the client and enabling server-side
    Redis caching. This architectural migration resolved the Wikidata
    soft-throttling limitation that had made interactive city search unreliable
    in the first version.

  \item \textbf{GeoNames city search pipeline.} A Python preprocessing script
    merges the GeoNames \texttt{allCountries.txt} and
    \texttt{alternateNamesV2.txt} datasets into a compact \texttt{cities.csv}
    index, which is imported into an Apache Solr~8 core with trigram
    tokenisation. The resulting pipeline supports fast, tolerant partial-word
    matching across 12~million populated places in English, Spanish, and
    Ukrainian, with population-based result ranking that can be tuned without
    modifying any external API.

  \item \textbf{Walter-Lieth climatogram engine.} A custom SVG rendering
    strategy using \texttt{clipPath} with the even-odd winding rule produces
    pixel-perfect arid and humid area fills without computing spline
    intersection points explicitly. Coordinate projection relies on
    Recharts~v3 internal hooks to make the overlay fully responsive to window
    resizing.

  \item \textbf{WorldClim pixel IRI parsing.} The heatmap engine parses the
    structured pixel IRIs returned by SCRAPI to reconstruct geographic tile
    bounds directly, eliminating a round-trip API call per pixel and enabling
    fast rendering of large regional selections.

  \item \textbf{Quality assurance pipeline.} A two-stage GitHub Actions CI/CD
    setup enforces TypeScript compilation, ESLint, Prettier, and unit test
    coverage on every commit, followed by a separate Playwright end-to-end
    suite that exercises the full request path including Redis and Solr.

  \item \textbf{Internationalisation.} The interface is delivered in three
    locales (English, Spanish, and Ukrainian) via \texttt{next-intl}, with
    city names returned by the GeoNames index in the active language.
\end{itemize}

% * ─── 5.2 Assessment of Objectives ────────────────────────────────────────────
\section{Assessment of Objectives}\label{CONC_OBJECTIVES}

Section~\ref{INTRO_OBJ} defined six specific objectives for the project.
Table~\ref{tab:objectives} assesses the degree to which each was achieved.

\begin{table}[H]
  \centering
  \caption{Assessment of project objectives.}
  \label{tab:objectives}
  \small
  \begin{tabular}{p{4cm}p{2cm}p{6.5cm}}
    \toprule
    \textbf{Objective} & \textbf{Status} & \textbf{Notes} \\
    \midrule
    WorldClim dataset access &
      \cmark~Met &
      All five climate normals and weather years 1951--2024 exposed through
      four interactive pages \\
    City climate profile &
      \cmark~Met &
      Standard climatogram and Walter-Lieth diagram implemented \\
    City comparison &
      \cmark~Met &
      Side-by-side comparison with overlaid charts and insight cards \\
    Period comparison &
      \cmark~Met &
      Climate mode (two normals) and weather mode (up to five years) \\
    Web architecture &
      \pmark~Revised &
      Original client-side-only objective superseded by full-stack Next.js
      architecture; rationale documented in Section~\ref{APP_ARCH_EVOLUTION} \\
    Internationalisation &
      \cmark~Met &
      English, Spanish, and Ukrainian delivered via \texttt{next-intl} \\
    \bottomrule
  \end{tabular}
  \normalsize
\end{table}

Five of the six objectives were fully met. The web architecture objective
underwent a deliberate revision during development: the original plan to build
a client-side-only single-page application was superseded by the need for
secure API key handling and controllable city search ranking, which required
a server-side component. This change was documented and justified in
Section~\ref{APP_ARCH_EVOLUTION} and ultimately produced a more robust and
maintainable architecture than the original design.

% * ─── 5.3 Evaluation Results ──────────────────────────────────────────────────
\section{Evaluation Results}\label{CONC_EVAL}

The usability evaluation conducted on 3 June 2026 with 27 students from ten
European universities produced encouraging results. The mean System Usability
Scale (SUS) score of \textbf{84.6} places Climatica in the \emph{Excellent}
range on the Bangor et al.\ adjective scale, well above the industry average
of 68. The Net Promoter Score of \textbf{+44} indicates that promoters
substantially outnumber detractors among the participant group.

The qualitative feedback aligned with the quantitative results. Ease of use
and intuitiveness were the most frequently mentioned positive attributes,
cited unprompted by approximately 80\% of respondents --- precisely the
audience for whom the application was designed. The two demo scenarios
developed in collaboration with Prof.\ Dr.\ Felipe Bravo (Catedrático de
Planificación Forestal, Universidad de Valladolid) demonstrated that Climatica
can support real forest management workflows --- from updating a 30-year
climate baseline for species selection to assembling evidence for climate-driven
forest decline --- in under twenty minutes, using only a web browser.

% * ─── 5.4 Lessons Learned ─────────────────────────────────────────────────────
\section{Lessons Learned}\label{CONC_LESSONS}

Several technical and methodological lessons emerged during the project.

\textbf{Prototype early, migrate deliberately.} The initial React SPA prototype
was essential for validating the SCRAPI integration and establishing the core
user interaction model before any architectural decisions were locked in. The
migration to Next.js was motivated by concrete, observable problems (Wikidata
throttling, uncontrollable ranking) rather than speculative concerns, which made
the decision easier to justify and the scope of the migration easier to bound.

\textbf{External API constraints shape architecture.} The two most significant
architectural decisions in the project --- moving city search from Wikidata to
Solr and adding a server-side caching layer --- were both driven by limitations
of external services rather than by internal design preferences. When building
on top of third-party APIs, it is worth investing early in understanding their
rate limits, ranking behaviour, and failure modes, as these constraints can
force substantial rework if discovered late.

\textbf{IRI structure as a performance optimisation.} The decision to parse
WorldClim pixel IRIs directly to reconstruct tile bounds, rather than issuing
a separate API call per pixel, eliminated a significant bottleneck in the
heatmap rendering pipeline. This approach was only possible because SCRAPI
follows the Linked Data principle of encoding resource properties in IRI
structure --- a concrete benefit of the semantic web architecture that might
not be immediately obvious to a developer approaching the API for the first time.

\textbf{Usability testing with domain users.} The evaluation session with
forestry and biology students highlighted issues --- particularly the manual
filter application step and heatmap loading times under fine resolutions ---
that were not apparent during development, when testing was performed by the
developer who already knew the application's interaction model. Domain-expert
users approach interfaces with different mental models, and their feedback
provided actionable insight that would have been difficult to obtain through
automated testing alone.

% * ─── 5.5 Future Work ─────────────────────────────────────────────────────────
\section{Future Work}\label{CONC_FUTURE}

The evaluation and development experience identified several concrete directions
for future work, which are summarised here and discussed in more detail in
Section~\ref{EVAL_DISCUSS_FUTURE}.

\textbf{Debounced filter workflow.} The most commonly requested usability
improvement was automatic filter application without an explicit ``Apply''
button. Redesigning the draft filter pattern to debounce API calls would
address this without reintroducing the problem of mid-edit partial states
triggering expensive requests.

\textbf{Raw data download.} Several technically sophisticated respondents
requested access to the underlying monthly pixel values. Adding a CSV or JSON
export of the full SCRAPI response would satisfy this request and extend
Climatica's utility to users who want to perform their own downstream analysis.

\textbf{Additional climate variables.} Climatica currently exposes only four
of the seven monthly variables available in the WorldClim climate dataset
(\texttt{tmin}, \texttt{tmax}, \texttt{tavg}, \texttt{prec}); solar radiation
(\texttt{srad}), wind speed (\texttt{wind}), and vapour pressure
(\texttt{vapr}) are not yet surfaced in the interface. The SCRAPI API also
exposes the full set of 19 WorldClim bioclimatic variables (BIO1--BIO19),
widely used in species distribution modelling and ecological niche analysis.
Exposing both the remaining monthly variables and the bioclimatic variables
would substantially extend Climatica's utility for researchers in biogeography
and ecology.

\textbf{Future climate projections.} WorldClim also distributes future climate
projections derived from global climate models (GCMs) under different emission
scenarios, covering periods such as 2021--2040 and 2041--2060. Integrating
these projections into Climatica --- alongside the historical data already
exposed --- would allow users to compare past trends with projected future
conditions, which is particularly relevant for the forest management use cases
described in Section~\ref{EVAL_DEMO}.

\textbf{Heatmap performance.} Two respondents reported slow loading on the
Region Heatmap page at fine resolutions. As discussed in
Section~\ref{EVAL_DISCUSS_FUTURE}, addressing this would first require
profiling whether the bottleneck lies in the SCRAPI API response time or in
client-side rendering.

\textbf{Longitude wraparound fix.} A known limitation identified during
development is that Leaflet allows the user to pan beyond the $\pm 180^{\circ}$
longitude boundary, producing coordinates that SCRAPI cannot process. Normalising
longitudes to $[-180^{\circ}, 180^{\circ}]$ before issuing API requests would
eliminate this edge case.

\textbf{Mobile touch support.} The map drawing interactions on the Region
Heatmap page are currently optimised for mouse input. Adding touch gesture
support would make this page fully usable on tablets and smartphones, addressing
feedback from iPad users encountered during the evaluation.

% ─── 5.6 Concluding Remarks ──────────────────────────────────────────────────
\section{Concluding Remarks}\label{CONC_REMARKS}

Climatica demonstrates that the WorldClim dataset --- previously accessible
only to GIS specialists and programmers --- can be made interactively explorable
for domain experts in forestry, ecology, and related fields through a carefully
designed web interface. The combination of the WorldClim Virtual Semantic API,
a locally hosted GeoNames search index, and a Next.js full-stack architecture
provides a foundation that is both performant and maintainable, and that can
be extended with additional features as the needs of its user community become
clearer.

The usability evaluation confirmed that the core design goal --- zero friction
for non-specialist users --- was successfully achieved. A mean SUS score of
84.6 and the qualitative feedback from forestry and biology students suggest
that Climatica occupies a genuine and previously unfilled position in the
landscape of climate data tools: between the static simplicity of
Climate-Charts.net and the expert-oriented power of NASA Giovanni or
Copernicus C3S.

The demo scenarios developed with Prof.\ Dr.\ Felipe Bravo illustrated that
Climatica is not merely a technical exercise but a practically useful tool
for real forest management decisions. Whether assessing the suitability of
tree species under a shifting climate normal or assembling spatial evidence
for bark beetle outbreak risk, the application provides actionable insights
in minutes rather than the hours or days that would be required with
traditional GIS workflows.

On a personal note, developing Climatica was a genuinely rewarding experience.
Knowing that the application would be used by real students and researchers ---
as confirmed by the evaluation session --- provided a constant source of
motivation throughout the project. The internship introduced me to technologies
and problem domains that were new to me as a developer: semantic web APIs,
geospatial data processing, and the specific challenges of building interactive
scientific tools for non-specialist audiences. I am grateful to my supervisors
and to Prof.\ Dr.\ Felipe Bravo for the opportunity to work on a project with
real practical value, and I hope that Climatica continues to be useful to the
forest management and ecology community at the Universidad de Valladolid and
beyond.